Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-04-14 manuscript/manuscript.tex
ms changes

Due to the high computational burden of forward simulations, we modeled a
single 50 megabase region in a population of $N = 1000$ diploid individuals
with a neutral variant mutation rate of $10^8$. This is roughly analogous to a
XXX chromosome of \emph{Drosophila melanogaster}; however with this small
population size and mutation rate, the population mutation rate $\theta$ for
the entire region leads to far fewer neutral sites to calculate covariances and
other statistics on than expected for a region this length in \emph{D.
melanogaster}. Since our main goal is to understand how the dynamics of
statistics used in the paper and how they are affected by background
selection, fitness model, and trait architecture, we use population
frequencies rather than sampling frequencies.

@@

\subsubsection{The Effects of the Genetic Architecture}

@@

@@

>>>>

Due to the high computational burden of forward simulations, we modeled a
single 50 megabase region in a population of $N = 1000$ diploid individuals
with a neutral variant mutation rate of $10^{-8}$ and a recombination rate of
$10^{-8}$ per basepair. This is roughly analogous to a XXX chromosome of
\emph{Drosophila melanogaster}; however with this small population size and
mutation rate, the population mutation rate $\theta$ for the entire region
leads to far fewer neutral sites to calculate covariances and other statistics
on than expected for a region this length in \emph{D.  melanogaster}. Since our
main goal is to understand how the dynamics of statistics used in the paper and
how they are affected by background selection, fitness model, and trait
architecture, we use population frequencies rather than sampling frequencies.

@@

\subsubsection{The Effects of the Genetic Architecture}
\label{supp:genetic-arch}

@@

In our analysis of the \textcite{Barghi2019-qy} data, we describe our statistic
$G(t)$ as an lower bound for two reasons: (1) the population is sequenced every
ten generations, meaning the temporal covariances between adjacent generations
cannot contribute to the numerator of $G(t)$ but contribute to the denominator,
and (2) the estimate of $G(t)$ ignores linked selection's contribution to the
variance in allele frequency change (this is the difference between $G$ and
$G'$, the latter of which includes these variance terms \cite{Buffalo2019-io}).
To verify that $G(t)$ estimated every ten generations is indeed a lower bound,
we used a simulation procedure similar to the exponential fitness function
simulations (described in Supplementary Material Section
\ref{supp:genetic-arch}), and calculated the temporal covariances and $G(t)$
both each generation, and every ten generations. Unlike the simulations
described in \ref{supp:genetic-arch}, we began selection at $10N$ generations,
and used trait $V_A = 0.01$ and $L = 1000$ sites affecting the trait.

First, comparing $G(t)$ when sampling population frequencies every generation
versus every ten generations, we confirm that the ten-generation block $G(t)$
is a lower bound of the true every generation $G(t)$ (red and blue lines in
Supplementary Figure \ref{suppfig:supp-blocks}A). Furthermore, since we control
the population size in our simulations at $N = 1000$ diploids, we know the
drift effective population size in the absence of selection. This allows us to
estimate $G(t)'$, which is a measure of $G(t)$ that accounts for the linked
selection's inflation of the variance in allele frequency change between two
generations (equation 26, \cite{Buffalo2019-io}). Plugging in the drift
effective population size $N_e = 1000$ into the expression for $G'(t)$ and
using the $\var(p_t - p_0)$ calculated for different $t$'s, we see that the
every generation $G(t)$ that does not account for linked selection's inflation
of $\var(\Delta p_t)$ does underestimate the true impact of linked selection as
expected (dashed gray line in Supplementary Figure \ref{suppfig:supp-blocks}A).

To further understand the effects of calculating temporal covariances every ten
generations rather than every generation, we also compared their magnitudes and
decay rates using the simulations described above. We find that ten generation
block temporal covariances are orders of magnitude larger but decay at similar
rates (see Supplementary Figure \ref{suppfig:supp-blocks}B; note the two y-axis
scales are different). The larger magnitude is expected, as each ten generation
block temporal covariance is the sum of

@@

  \caption{A: The $G(t)$ averaged over 50 replicate simulations with $V_A =
    0.01$ and $L = 1000$. The blue line shows $G(t)$ calculated over ten
    generation blocks, similar to the calculation of temporal covariances of
    the \textcite{Barghi2019-qy} study. The red line shows the average $G(t)$
    estimates when the population is sampled every generation and all
    covariances can contribute to the numerator of $G(t)$. The dashed gray line
    indicates the $G(t)'$ estimate, which uses the known drift effective
    population size of the simulations. B: The temporal covariances calculated
    each generation (red line) and on ten generation blocks (blue line) using the
    same simulation data.}
